---
title: "National Parks"
author: "Evan Canfield"
date: "6/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Library Calls
```{r library calls, message=FALSE, warning=FALSE}
if (!require(tidyverse)) {install.packages('tidyverse')} 
library(tidyverse)
if (!require(readr)) {install.packages('readr')} 
library(readr)
if (!require(stringr)) {install.packages('stringr')} 
library(stringr)
if (!require(data.table)) {install.packages('data.table')} 
library(data.table)
if (!require(broom)) {install.packages('broom')} 
library(broom)
if (!require(plotly)) {install.packages('plotly')} 
library(plotly)
if (!require(ggthemes)) {install.packages('ggthemes')} 
library(ggthemes)
if (!require(RColorBrewer)) {install.packages('RColorBrewer')} 
library(RColorBrewer)
```

#Data Import / Export
```{r data import,  message=FALSE,}
#Visitation Data
nps_summary_all<- read.csv("./data/annual_summary_report_1904-2018.csv", stringsAsFactors = FALSE)

#Location and Region Data on the Parks
list_np_locations <- read.csv(file = "./data/list_national_parks_location.csv", stringsAsFactors = F)

#Import NPS Region Color Table
nps_region_colors <- read.csv("./data/nps_region_colors.csv", stringsAsFactors = FALSE)
```

#Color Map 
```{r}
#Assign Colors to NPS Regions
color_map <- setNames(nps_region_colors$html_color_code_stamps,nps_region_colors$nps_region)
```


#Data Inspection
```{r}
glimpse(nps_summary_all)

nps_rec_visits_all <- nps_summary_all %>% 
  select(ParkName, Year, RecreationVisitors)

glimpse(nps_rec_visits_all)
```

#Reduce to National Parks
```{r}
#Add NP to American Samoa for Downstream Filters
#Remove National Park from the name. This is a formatting choice.
nps_rec_visits_all$ParkName <- str_replace_all(nps_rec_visits_all$ParkName, "National Park of American Samoa", "American Samoa NP")

nps_rec_visits_all %>% 
  distinct(ParkName)

#Filter for National Park (NP) and National Preserve (NPRES)
#This is due to some national parks also being National Preserves
nps_rec_visits_np <- nps_rec_visits_all %>% 
  filter(str_detect(string = ParkName, " NP")) %>% 
  filter(str_detect(string = ParkName, " NPRES") == FALSE)

#Generate distinct list of national parks
nps_rec_visits_np %>% 
  distinct(ParkName)

#Join Location Data by ParkName
nps_rec_visits_np <- nps_rec_visits_np %>% 
  left_join(y = list_np_locations, by = c("ParkName"))



```

#Line Graph - All Values
```{r}
l <- ggplot(data = nps_rec_visits_np, 
            mapping = aes(x = Year, y = RecreationVisitors, color = ParkName)) +
  geom_line() +
  guides(color = FALSE)

ggplotly(l)
```

#Drop Years from Before Entity Became a National Park
```{r}
nps_rec_visits_np_only <- nps_rec_visits_np %>% 
  filter(Year >= np_established_year )

nps_rec_visits_np_only

m <- ggplot(data = nps_rec_visits_np_only, 
            mapping = aes(x = Year, y = RecreationVisitors, color = ParkName)) +
  geom_line() +
  guides(color = FALSE)

ggplotly(m)

```

#Identify Top 10
```{r}
glimpse(nps_rec_visits_np)

year_filter = 1800

nps_top_10 <- nps_rec_visits_np %>% 
  filter(Year >= year_filter) %>% 
  group_by(Year) %>%
  mutate(total_parks = n_distinct(ParkName)) %>% 
  mutate(visitors_rank = rank(RecreationVisitors)) %>% 
  mutate(np_rank = as.integer(total_parks - visitors_rank + 1)) %>% 
  ungroup() %>% 
  select(ParkName, np_id, np_rank) %>% 
  filter(np_rank <= 10) %>% 
  distinct(ParkName, np_id)

nps_top_10
```

#Line Graph Top 10
```{r}
nps_rec_visits_np_top_10 <- nps_rec_visits_np %>% 
  semi_join(y = nps_top_10, by = c("np_id"))

nps_rec_visits_np_top_10

l <- ggplot(data = nps_rec_visits_np_top_10, 
            mapping = aes(x = Year, y = RecreationVisitors, color = ParkName)) +
  geom_line() +
  theme_minimal()

ggplotly(l) %>% 
  layout(legend = list(x = 0.05, y = 0.99))
```

#Data Prep - Circular Bar Chart - R Grpah Gallery 295 - 
```{r}
#Filter for 2018
nps_rec_visits_np_2018 <- nps_rec_visits_np %>% 
  filter(Year == 2018)

#Filter Down to Only Park Name
nps_rec_visits_np_2018$ParkName <- str_remove_all(nps_rec_visits_np_2018$ParkName," NP")

glimpse(nps_rec_visits_np_2018)

nps_rec_visits_np_2018$ParkName <- str_remove_all(nps_rec_visits_np_2018$ParkName," & PRES")

glimpse(nps_rec_visits_np_2018)

#Filter Out Wolf Trap
nps_rec_visits_np_2018 <- nps_rec_visits_np_2018 %>% 
  filter(ParkName != 'Wolf Trap for the Performing Arts')

nps_rec_visits_np_2018 <- nps_rec_visits_np_2018 %>% 
  mutate(percent_visit = round(RecreationVisitors/sum(RecreationVisitors)*100,digits = 2))

glimpse(nps_rec_visits_np_2018)

nps_rec_visits_np_2018 %>% 
  arrange(np_id)
```

#Circular Bar Chart - R Grpah Gallery 295
```{r}
p_295 <-  
  ggplot(data = nps_rec_visits_np_2018, 
          mapping = aes(x = as.factor(np_id), y = percent_visit)) +
  geom_bar(stat = "identity", fill = alpha("blue", 0.3)) +
  ylim(-5, 13.5) +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = unit(rep(-2,4), "cm")     # This remove unnecessary margin around plot
  ) 

p_295

#Wrap in Polar Coordinates
p_295  + coord_polar(start = 0)

```

#Circular Bar Chart With Labels - R Grpah Gallery 296
```{r}
labels <- nps_rec_visits_np_2018
 
# calculate the ANGLE of the labels
number_of_bar <- nrow(labels)

angle <- 90 - 360 * (labels$np_id - 0.5) / number_of_bar     

# I substract 0.5 because the letter must have the angle of the center of the bars. 
#Not extreme right(1) or extreme left (0)
 
# calculate the alignment of labels: right or left
# If I am on the left part of the plot, my labels have currently an angle < -90
labels$hjust <- ifelse( angle < -90, 1, 0)
 
# flip angle BY to make them readable
labels$angle <- ifelse(angle < -90, angle + 180, angle)

#plot
p_296 <-  
  ggplot(data = nps_rec_visits_np_2018, 
          mapping = aes(x = as.factor(np_id), 
                        y = percent_visit)) +
  geom_bar(stat = "identity", fill = alpha("blue", 0.3)) +
  ylim(-3, 13.8) +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = unit(rep(-1,4), "cm")     # This remove unnecessary margin around plot
  ) +
  coord_polar(start = 0) +
  geom_text(data = labels, 
            mapping = aes(x = np_id, 
                          y = percent_visit + 0.5, 
                          label = ParkName,
                          hjust = hjust), 
            color="black", 
            fontface="bold", 
            alpha=0.6, 
            size=2, 
            angle= labels$angle, 
            inherit.aes = FALSE)

p_296
```

#Circular Bar Chart With Labels, Seperated By Region - R Grpah Gallery 296
```{r}
nps_rec_visits_np_2018_cir <- nps_rec_visits_np_2018

#Make NPS Region a Factor
nps_rec_visits_np_2018_cir$nps_region <- as.factor(nps_rec_visits_np_2018_cir$nps_region)

#Arrange Data Input
 nps_rec_visits_np_2018_cir <- nps_rec_visits_np_2018_cir %>%
  arrange(nps_region, RecreationVisitors)

# Set a number of 'empty bar' to add at the end of each group
empty_bar <- 3
to_add <- data.frame( matrix(NA, empty_bar*nlevels(nps_rec_visits_np_2018_cir$nps_region), ncol(nps_rec_visits_np_2018_cir)))
colnames(to_add) <- colnames(nps_rec_visits_np_2018_cir)
to_add$nps_region <- rep(levels(nps_rec_visits_np_2018_cir$nps_region), each=empty_bar)
nps_rec_visits_np_2018_cir <- rbind(nps_rec_visits_np_2018_cir, to_add)
nps_rec_visits_np_2018_cir <- nps_rec_visits_np_2018_cir %>% arrange(nps_region)
nps_rec_visits_np_2018_cir$np_id <- seq(1, nrow(nps_rec_visits_np_2018_cir))


#Assign Labels to the Bars
labels <- nps_rec_visits_np_2018_cir
 
# calculate the ANGLE of the labels
number_of_bar <- nrow(labels)

angle <- 90 - 360 * (labels$np_id - 0.5) / number_of_bar     

# I substract 0.5 because the letter must have the angle of the center of the bars. 
#Not extreme right(1) or extreme left (0)
 
# calculate the alignment of labels: right or left
# If I am on the left part of the plot, my labels have currently an angle < -90
labels$hjust <- ifelse( angle < -90, 1, 0)
 
# flip angle BY to make them readable
labels$angle <- ifelse(angle < -90, angle + 180, angle)

#plot
p_297 <-  
  ggplot(data = nps_rec_visits_np_2018_cir, 
          mapping = aes(x = as.factor(np_id), 
                        y = percent_visit,
                        fill = nps_region)) +
  geom_bar(stat = "identity", alpha = 0.5) +
  ylim(-3, 13.8) +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = margin(0, 0, 0, 0, "cm"), # This remove unnecessary margin around plot
    legend.position = "none"
    ) +
  coord_polar(start = 0) +
  geom_text(data = labels,
            mapping = aes(x = np_id,
                          y = percent_visit + 0.5,
                          label = ParkName,
                          hjust = hjust),
            color="black",
            fontface="bold",
            alpha=0.6,
            size=2,
            angle= labels$angle,
            inherit.aes = FALSE)

p_297
```

#Plotly - Circle
```{r}
nps_rec_visits_np_2018_py <- nps_rec_visits_np_2018 %>% 
  arrange(RecreationVisitors) %>% 
  mutate(RecreationVisitors.fm = prettyNum(RecreationVisitors, big.mark = "," , scientific = FALSE, ))

glimpse(nps_rec_visits_np_2018_py)

# Get incremental angle value

open_space <- 0

n <- nrow(nps_rec_visits_np_2018_py) + open_space

dtheta <- 2*pi / (n-1)

theta <- pi / 2
 
# Initialise
l <- c()
x <- c()
y <- c()
xend <- c()
yend <- c()


# Radius of the center white-space circle. If set to zero all bars would eminate from the same point when plotting a circle.
adjust <- 2

# Calculate x and y coordinates
for(ctr in 1:nrow(nps_rec_visits_np_2018_py)){
  
  l[ctr] <- nps_rec_visits_np_2018_py$percent_visit[ctr] + adjust
  
  x[ctr] <- adjust * cos(theta)
  y[ctr] <- adjust * sin(theta)
  
  xend[ctr] <- l[ctr] * cos(theta)
  yend[ctr] <- l[ctr] * sin(theta)
  
  theta <- theta + dtheta
}

l
x
y
xend
yend
dtheta

plot.df <- data.frame(x, y, xend, yend, nps_region = nps_rec_visits_np_2018_py$nps_region)

plot.df

p <- plot_ly(plot.df, 
             x = ~x, y = ~y,
             xend = ~xend, yend = ~yend,
             color = ~nps_region) %>% 
  add_segments(line = list(width = 5),
               hoverinfo = 'text',
               text = ~paste(nps_rec_visits_np_2018_py$ParkName,
                      '</br> </br> Visitors: ', nps_rec_visits_np_2018_py$RecreationVisitors.fm)
                      )

p
```

#Plotly - sprial, Create Data Frames
```{r}
# #List of National Parks By Percent Visited. 
# #This will be used to set the factor levels of the National Park Names
#  np_order <- nps_rec_visits_np_2018 %>%
#    arrange(percent_visit) %>%
#    select(ParkName) %>%
#    #Convert df to list
#    pull()
# 
# #Reset the Park Name Factor levels to be ordered by percent visit.
# #If the levels of this factor are not manually set,  Plotly will default the segemnts by alphabetical order by Park Name
# nps_rec_visits_np_2018$ParkName <- factor(nps_rec_visits_np_2018$ParkName, levels = np_order)

#Format Number Type. Convcert Number of visitors to a number with comma seperators
nps_rec_visits_np_2018_py <- nps_rec_visits_np_2018 %>% 
  mutate(RecreationVisitors.fm = prettyNum(RecreationVisitors, big.mark = "," , scientific = FALSE, )) %>% 
  arrange(percent_visit)
```

#Plotly - sprial, Initial Plot Layout
```{r}
# Adds additional spaces in the data table to create offset if open_space > 0
open_space <- 10

#Number of spaces within the spiral.
#If open space is set to zero the bar chart will fully rotate around the axis
n <- nrow(nps_rec_visits_np_2018_py) + open_space

#Attributes for determining placement around axis
dtheta <- 2*pi / (n-1)

theta <- pi * (1/2)
 
# Initialise For Loop Entities
l <- c()
x <- c()
y <- c()
id <- c()
xend <- c()
yend <- c()
segment_len <- c()

#Spiral Equaltion Constants
A <- 0.25
B <- 0.25

# Radius of the center white-space circle. If set to zero all bars would eminate from the same point when plotting a circle.
adjust <- 5

# Calculate x and y coordinates for bars
for(ctr in 1:nrow(nps_rec_visits_np_2018_py)){
  
  id[ctr] <-nps_rec_visits_np_2018_py$np_id[ctr]
  l[ctr] <- nps_rec_visits_np_2018_py$percent_visit[ctr] + adjust
  
  x.eq <- A*cos(theta)*exp(B*theta) 
  y.eq <- A*sin(theta)*exp(B*theta) 
  
  x[ctr] <- adjust * x.eq
  y[ctr] <- adjust * y.eq
  
  xend[ctr] <- l[ctr] * x.eq
  yend[ctr] <- l[ctr] * y.eq
  
  theta <- theta + dtheta
}

#Generate Dataframe for Plotly visual
plot.df <- data.frame(id, x, y, xend, yend)
```

With the dataframe of plot points now created, we now join that raw data to the national parks dataframe and plot the visual
```{r}
#Join New Dataframe with National Park Visitor Information Dataframe and Color Dataframe
plot.df_p1 <- plot.df %>% 
  left_join(y = nps_rec_visits_np_2018_py, by = c("id" = "np_id"))

glimpse(plot.df_p1)

#Bar Chart is arranged by aplphabetic order. We need to incorporate factor levels to force the bar chart
# to ascending order by percent visited
p1 <- plot_ly(plot.df_p1,
              name = ~nps_region) %>% 
  add_segments(x = ~x, xend = ~xend,
               y = ~y, yend = ~yend,
               color = ~nps_region,
               colors = color_map,
               line = list(
                 width = 6), 
                 hoverinfo = 'text',
                 text = ~paste(ParkName,
                       '</br> </br> Visitors: ', RecreationVisitors.fm)
                       )
p1
```

#Add Formatting
```{r}
#Add Legend Format and Erase X and Y Axis marks
legend_layout <- list(
  name = ~nps_region,
  font = list(
    family = "sans-serif",
    size = 14,
    color = "#000"),
  bg = "transparent",
  #orientation = "h",
  x = 0, 
  y = 1)


p2 <- layout(p1,
            xaxis = list(title = "", showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE, range = c(-4,12)),
            yaxis = list(title = "", showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
            legend = legend_layout,
            autosize = TRUE
            #width = 500, height = 800
)

p2
```

#Add Labels
```{r}
#Add National Park Names, Calculate Park Name locations
textangle <- 90
 
for(ctr in 1:nrow(nps_rec_visits_np_2018_py)){
  
  l[ctr] <- nps_rec_visits_np_2018_py$percent_visit[ctr] + adjust

  x.eq <- A*cos(theta)*exp(B*theta) 
  y.eq <- A*sin(theta)*exp(B*theta) 
  
  x[ctr] <- adjust * x.eq
  y[ctr] <- adjust * y.eq
  
  xend[ctr] <- l[ctr] * x.eq
  yend[ctr] <- l[ctr] * y.eq
  
  x_len[ctr] <- xend[ctr] - x[ctr]
  y_len[ctr] <- yend[ctr] - y[ctr]
  
  segment_len[ctr] <- sqrt((xend[ctr]-x[ctr])^2 + (yend[ctr]-y[ctr])^2)

  
    if(ctr < 30) {xanchor <- "right"; yanchor <- "middle"}
    if(ctr > 31 & ctr < 50) {xanchor <- "right"; yanchor <- "middle"}
    if(ctr > 51) {xanchor <- "left"; yanchor <- "middle"}


  p2$x$layout$annotations[[ctr]] <- list(x = xend[ctr], y = yend[ctr], showarrow = FALSE,
                                        text = paste0(nps_rec_visits_np_2018_py$ParkName[ctr]),
                                        textangle = textangle,
                                        xanchor = xanchor,
                                        yanchor = yanchor,
                                        font = list(family = "sans-serif", size = 10),
                                        borderpad = 0,
                                        borderwidth = 0)
  theta <- theta + dtheta
  textangle <- textangle - (180 / pi * dtheta)
  
  if(textangle < -90) textangle <- 90
}

p3 <- p2

plot.df_p3 <- data.frame(x, y, xend, yend, x_len, y_len, segment_len)

plot.df_p3

p3
```